---
title: "Data Import"
output: github_document
---


```{r serup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data.

### Absolute path       
Get current working directory using getwd()    
Disadvantages: 1. too complex. 
2. Only work in your own computer
### Relative path
### Import the FAS litters CSV
```{r}
litters_df = read_csv("data/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)


#Lowercases all column names: It converts all column names to lowercase for uniformity.
#Replaces spaces and special characters: Spaces and non-alphanumeric characters (e.g., !, @, #) are replaced with underscores (_).
#Removes leading and trailing whitespace: Any leading or trailing whitespace in column names is removed.
#Ensures names are unique: If there are duplicate column names, it appends a numeric suffix to ensure that each column has a unique name.'''

#head(litters_df)
#tail(litters_df)
```

```{r eval=FALSE}
View(litters_df)
```

```{r, message=FALSE}
# relative path
pups_df = read_csv("data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)

# Absolute path
pups_df = read_csv("/Users/leyangrui/Desktop/Columbia/Data Science/data_wraggling1/data/FAS_pups.csv")
# --> very fragile, does not work if you move files around
```
### Look at read_csv options

col names and skipping rows

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    #skip = 2
  )
```

Missing data
and data that interupts data type

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )

litters_df = janitor::clean_names(litters_df)

gd0_weight = pull(litters_df, gd0_weight)
mean(gd0_weight, na.rm = TRUE)
```

What if we want to code 'group' as a factor variable?

```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )
```

### Import an excel file

Import MLB 2011 summary data

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```

### Import SAS data

```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```

## Never use read.csv()

```{r}
# litters_df = read.csv("data/FAS_litters.csv")
```
read in without tibble reminders

## Never use $ to pull out variables
Do not take variables out from a data frame.











